Unifying quasars with clumpy wind models

### 2015-07-30 paper_mnras.tex
nearly there for first draft

\maketitle
\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that clumping enhances the H recombination and collisionally excited resonance lines,
causing strong line emission (EW=?) to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to produce line emission with comparable equivalent widths
to existing quasar composites, due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. We briefly 
explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss a number of potential solutions. 
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next steps if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/uvspec.eps}
\caption
{
Synthetic spectra at four viewing angles in the clumpy model. Plot would look different probably,
and may show composites, but this would be the main synthetic spectrum plot.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/opt.eps}
\caption
{
Figure designed to show optical spectrum and comparison to composite.
}
\label{fig:uvspec}
\end{figure*} %fullpage

Figure 4 shows the synthetic spectrum in the UV from the two models, with the
benchmark model from H13 also shown for comparison. We also show a comparison
to composite quasar and BALQSO spectra in figure 5, with the optical waveband 
included. In the UV, strong BAL profiles are formed in both models. This
highlights the first success of our model: clumping allows the correct ionization state 
to be maintained in the presence of strong X-rays. 
The BAL profiles are saturated, 
and generally show the strongest absorption at the terminal velocity.
This contradicts observed profiles of many BALQSOs (REFs), 
although there are some objects which do show this behaviour (REFs). 
However, we have found that this profile shape is by no means ubiquitous.
Figure 6 shows how the line profile varies with angle for three of the UV lines
(see also section 4.4). In addition, the line profile
shape is strongly dependent on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the line profile shape. Figure ?? shows selection of line profiles 
for different kinematic parameters to demonstrate this. 
The degeneracy between parameters, and lack of available
physical and observational constraints goes someway to showing why individual 
fitting of line profile shapes with kinematic models is such a difficult
exercise.

We find that model A can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. However, the red wing of the BAL profiles 
is far stronger than seen in BALQSO spectra and composites.
This is due to a combination of strong collisionally excited emission
and continuum attenuation. To illustrate the amount of 
continuum attenuation we also show the continuum from a disc-only model.
The line-to-continuum ratios at low inclinations are also
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line equivalent widths 
(EWs) -- this is discussed further in section~5.

To assess the ability of the model to match real 
quasar spectra, we show a comparison to an HST quasar composite in figure 5. 
All the spectra are normalised to the flux at 2000\AA in these plots.
Model A matches the \civ\ line profile well, but the \la\ profile
is a factor of $\sim3$ weaker than in the composite spectrum.
Figure 5 also shows the spectrum into the optical,
showing that the outflow produces signiciant optical emission lines
as well as having a noticeable effect on the continuum shape. This
is particularly true around the Balmer jump at $3646\AA$ and is due to
a prominent wind-formed recombination continuum.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/alpha_ox_both.eps}
\caption
{
X-ray properties of the H13 and clumped model (text filled 
squares), plotted against monochromatic luminosity 
at 2500\AA. Also plotted are the samples considered by
Saez et al. 2012 on a similar plot; The COMBI-7 AGN sample (ref),
the BQS sample (ref) and the Saez et al. (2012) sample of BALQSOs.
}
\label{fig:xray}
\end{figure*} %fullpage

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/sed_all_balqsos.eps}
\caption
{
Broadband SEDs compared to IR and X-ray SEDs for selected BALQSOs 
from Grupe \& Nousek (2015).
}
\label{fig:xray}
\end{figure*}

The bottom panel of figure~\ref{fig:xray} shows the emergent $L_{2keV}$ 
monochromatic luminosity ($L_\nu$) and $\alpha_{OX}$ plotted against 
$L_\nu$ at $2500$\AA\ for a number of different viewing angles in our model.
$\alpha_{OX}$ is a spectral index from near UV to X-rays defined by

\begin{equation}
\alpha_{OX}=0.3838\log\left(\frac{L_{\nu}(2~keV)}{L_{\nu}(2500~\mbox{\scriptsize{\AA}})}\right).
\end{equation}

The properties are calculated from the synthetic spectra and thus include
the effects of wind reprocessing and attenuation. In addition to model outputs,
we also show the BALQSO sample of Saez et al. (2012) and luminous AGN and quasar
samples from Steffen et al. (2006). For low inclination, `quasar-like' viewing angles,
we now show excellent agreement with AGN samples. The trend with inclination
in our models is caused by a combination of disc foreshortening/limb-darkening 
(resulting in a lower $L_{2500}$ for higher inclinations) and the fact that the disk 
is opaque and thus the X-ray source subtends a smaller solid angle at high inclinations
(resulting in a lower $L_{2keV}$ for higher inclinations).

Although the low inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
it appears that our models tend to underpredict the emergent X-ray luminosity at 
although we are limited by poor statistics. 
It is possible that this is due to our wind being overly optically thick to 
electron scattering at angles which look through the wind. It could also be that the
isotropic X-ray source assumption is incorrect and BALQSOs are {\em intrinsically} 
X-ray weak \citep[e.g.][]{morabito2013}. Nevertheless, our input X-ray spectrum
now reproduces the X-ray properties of a luminous quasar.

Typically, BALQSOs show strong X-ray absorption with columns 
of $N_H\sim10^{23}~\rm{cm^{-2}}$ 
\citep{green1996,brandt2000,mathur2000,green2001,grupemathur2003}.
This is often cited as evidence that the BAL outflow is shielded from
the X-ray source, especially as sources with strong X-ray absorption tend
to exhibit deep BAL troughs and high outflow velocities (REFs).
Our results tend to imply that a clumpy BAL outflow
itself can be responsible for the strong X-ray absorption, and possibly 
explains why mini-BALs have weaker X-ray absorption than BALQSOs 
\citep[e.g.][]{brandt2000,hamann2013}.

We can also examine the emergent X-ray spectra from our model. 
Our code is not yet optimized for producing X-ray spectra, as we do not yet 
include all lines and levels of e.g. Fe or macro-atom treatments of 
higher order ions (cf. Sim et al. 2008, 2010). 
However, the bound-free opacity sources are complete, and comparisons of the shape
of the heavily absorbed spectra can be useful.
Figure 6 shows the broadband SED from the model, compared to a number of BALQSO spectra
from \cite{grupenousek2015} - note that lower mass and super-Eddington sources
are excluded.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/c4_with_i.eps}
\caption
{
Carbon IV, Al III and Mg II line profiles with inclination for wind angles.
Showing how the line profiles change
and how LoBALs kick in at particularly high inclinations.
}
\label{fig:lobal}
\end{figure*} %fullpage

At certain sightlines, our model now produces blue-shifted BALs in \al\ and \mg --
the absorption lines seen in LoBALQSOs. Line profiles in velocity space 
for \civ, \al\ and \mg, are shown in figure~\ref{fig:lobal} for a range
of BALQSO viewing angles. We confirm the behaviour expected from 
a unification model such as Elvis (2000), in which ionization stratification
explains the incidence of LoBALQSOs. In this framework, the lower
ionization material is intercepted by a smaller family of sightlines.
This is also the explanation here, as shown by figure~\ref{fig:wind}.

These line profiles also indicate a potential problem with our model. 
\cite{odowd2015} find that the LoBAL profiles in H1413+117 have a 
higher velocity onset than the higher ionization BALs, suggesting that the
wind becomes less ionized as radius increases. 
We find the opposite trend in velocity onsets. In addition, examination of 
figure~\ref{fig:wind} shows that ionization parameter decreases with radius, 
because the decrease in density wins over geometric dilution of the radiation 
field. While the \cite{odowd2015} hypothesis comes from one object,
it is clear that {\em at least some} BALQSOs must become less ionized at 
higher velocities, contrary to what one expects from our modelling.

@@

\clearpage
\section{Discussion: The Disc SED}

\subsection{Anisotropy of disc emission}

Discuss the importance of the angular distribution of the disc SED on line
(limb-darkened, foreshortened, etc.)

@@

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at $2000$~\AA\ as a function of inclination
for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{f2000}.
The effect of relativistic effects on disc anisotropy can be clearly seen.

To examine the effect of GR on our models, we carry out the following procedure.
First, we fit our output spectrum with a polynomial. This continuum fit can 
be subtracted from the actual spectrum 
to produce a model of the line emission only. We then calculate the 
`wind continuum factor', by comparing the model fit with the emergent 
disc spectrum without a wind. This factor gives us an approximate way to quantify the 
effect of the wind on a disc continuum. It will generally be $>1$ at high inclinations,
and $<1$ at low inclinations. This calculates emergent accretion disc spectra from an AGN disc atmosphere
including relativistic effects. Finally, produce our final synthetic spectrum 
by convolving this the line emission spectrum with the \agn\ model for the model parameters.

Figure 10 shows the resultant spectrum. Including GR clearly helps

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/f2000_m9.eps}
\caption
{
$F_{2000}$\ as a function of inclination from
\agn\ models, compared to a classical AD. 
{\bf This is a placeholder- it will show the effect of GR on the anisotropy
of the disk for different wavelengths and eddington fractions, compared
to limb darkened and foreshortened classical AD}.
}
\label{fig:f2000}
\end{figure}

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/adjusted_withbals.eps}
\caption
{
$F_\lambda$ normalised to $F_{2000}$. {\bf Again, this is a placeholder- but I'm thinking some kind of comparison to composite including the adjusted continuum, showing that we can't get it exactly right}.
}
\label{fig:finalcomp}
\end{figure*} %fullpage

How would wind reprocessing help?

\subsection{The BALQSO fraction and wind covering factor}

The disc SED has a profound effect on the intrinsic BAL fraction
inferred from flux-limited samples. This in turn impacts on 
the geometry and covering factor assumed for BALQSO and unified wind models,
and even effects the efficiency of feedback.
\cite{krolik1998} noted that the effect of limb darkening and foreshortening
would result in a signicantly underestimated covering factor for BALQSO winds.
The effect is complicated further by the quasar luminosity function -- if 
BALQSOs have lower observed luminosities for given accretion disc properties 
then they may be drawn from a different bin in the quasar luminosity function 
\citep{goodrich1997}. These effects are understandably not taken into account
by current estimates of the intrinsic BALQSO fraction \citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.
% These factors will also combine in the building-up 
% of BALQSO composite spectra, and mean that is very difficult to 
% disentangle the true nature of BALQSOs
% without BH mass and eddington ratio estimates.
The way forward is somewhat unclear. 
Given the lack of available observational constraints, models cannot readily 
be used to inform the selection criteria when calculating the BALQSO fraction.
Nevertheless, understanding the true nature of the disc SED, including the
angular distribution of radiation, is clearly crucial in order
to accurately estimate the true covering factor of BALQSO winds.

@@

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/f2000_m9.eps}
% \caption
% {
% $F_\lambda$ at $2000$~\AA\ as a function of inclination from
% \agn\ models. Spectra are computed for BH of $10^9~M_\odot$ with
% a number of different BH spins and Eddington fractions. The black line
% shows a standard multi-temperature blackbody AD model.
% }
% \label{fig:alpha_ox}
% \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Summary}

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We have shown that the degree of ionization stratification
in the model is sufficient that LoBAL line profiles
are seen at a subset of viewing angles. 
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models are not comparable to those in Quasar composite
spectra. This is due to a fundamental constraint discussed in section ?. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry. 
\smallskip
\item We have examined the effect of GR on our disc SED, using the disc atmosphere
and GR ray-tracing code \agn. While including GR effects
does cause the disc SED to become significantly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We discuss other solutions to this problem in section ?; 
It is possible that a combination of GR and reprocessing by the wind could provide a 
solution, and a number of complicated selection effects may be at work
in the building of the quasar composites.
\end{enumerate}

Our work confirms a number of expected outcomes from such a model, and suggests 
that a simple geometry such as this can come close to explaining much of the 
phenomenology of quasars. Nevertheless, our conclusions pose a clear challenge 
to the current unification picture.

>>>>

\maketitle
\begin{abstract} 
% Broad absorption lines (BALs) in the ultraviolet 
% are seen in $\sim20\%$ of quasi-stellar objects (QSOs). 
% Blue-shifted broad absorption lines (BALs) are the most direct evidence of 
% accretion disc `winds' in such systems; mass loaded outflows
% emanating from the disc that may be driven by line forces or
% magnetic processes. 
Various unification schemes for
quasars and luminous active galactic nuclei (AGN) have proposed
that the broad emission line region is roughly cospatial
with broad absorption line (BAL) gas and much of the phenomenology of luminous AGN
can be explained by a simple geometrical picture involving an accretion
disc and associated outflow. Here, we test this paradigm by 
utilising our state-of-the-art radiative transfer code to produce synthetic spectra
from simple biconical disc wind models. In particular, we expand on our previous 
work in which a benchmark model for BAL quasars was produced. 
We have conducted a limited parameter search with the aim of unifying quasar phenomenology,
and arrive at an improved model. The grid is now publicly available.
We find that a simple treatment of clumping (`microclumping') 
allows for a more realistic X-ray luminosity in the model by lowering the 
ionization parameter. We examine the X-ray properties of this new model
and find good agreement with existing X-ray samples of AGN and QSOs.
We find that the dense, X-ray heated wind 
produces strong H recombination and collisionally excited resonance 
line emission to emerge at the low inclination angles, 
which represent quasars within this unification scenario.
% We also treat Hydrogen using a `macro-atom' approach in order to 
% examine the effect of recombination on Hydrogen emission lines, and
% this results in significant line emission in Lyman~$\alpha$ and the Balmer series.
However, we are unable to reproduce the remarkably similar of line-to-continuum
ratios between BAL and non-BAL quasar composites.
This is due to a fundamental constraint arising
from the anisotropy of emission from a classical thin disc. 
We briefly explore the effect of relativistic beaming, gravitational redshift and 
light bending on the angular distribution of disc continuum emission. 
We find that these general relativistic effects do cause the disc to
emit slightly more isotropically, but this is not yet sufficient to produce a
self-consistent model. We discuss wind reprocessing as a potential solution.
Overall, our work suggests that geometric unification
involving an accretion disc wind is a promising scenario, but our results 
pose a number of difficult challenges to such a model.
% Determining the true geometry of ADWs and uncovering the true disc spectral 
% (and angular) energy distribution are key next stpng if we are to build up a 
% holistic picture of the quasar population.
\end{abstract}

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec.png}
\caption
{
Synthetic spectra at four viewing angles in the clumpy model. Plot would look different probably,
and may show composites, but this would be the main synthetic spectrum plot.
}
\label{fig:uvspec}
\end{figure*} %fullpage

% \begin{figure*} %fullpage
% \centering
% \includegraphics[width=1.0\textwidth]{figures/opt.png}
% \caption
% {
% Figure designed to show optical spectrum and comparison to composite.
% }
% \label{fig:uvspec}
% \end{figure*} %fullpage

Figure 4 shows the synthetic spectrum in the UV from the two models, with the
benchmark model from H13 also shown for comparison. We also show a comparison
to composite quasar and BALQSO spectra in figure 5, with the optical waveband 
included. In the UV, strong BAL profiles are formed in both models. This
highlights the first success of our model: clumping allows the correct ionization state 
to be maintained in the presence of strong X-rays. 
The BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At lower inclinations, the strongest absorption occurs at the red edge,
whereas at high inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.
% the line profile shape. Figure ?? shows selection of line profiles 
% for different kinematic parameters to demonstrate this. 
% The degeneracy between parameters, and lack of available
% physical and observational constraints goes someway to showing why individual 
% fitting of line profile shapes with kinematic models is such a difficult
% exercise.

We find that the model can produce significant line emission
at low inclinations, particular in \civ, 
and the improved treatment of recombination results in a strong \la\
line. In the context of unification, this is a promising results, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles.
However, the red wing of the BAL profiles 
is generally stronger than seen in BALQSO spectra and composites.
This is due to a combination of strong collisionally excited emission
and continuum attenuation. To illustrate the amount of 
continuum attenuation we also show the continuum from a disc-only model.
The line-to-continuum ratios at low inclinations are also
significantly affected by disc foreshortening and limb darkening.
The angular distribution of the disc radiation is clearly
crucially important in determining the emergent line equivalent widths 
(EWs) -- this is discussed further in section~5.

To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
The model matches the \civ\ and \mg\ line profiles fairly well, 
but the \la\ profile is a factor of $\sim3$ weaker than in the composite spectrum.
This is implies that a larger recombination line formation region is required in order 
to successfully reproduce the strong \la\ emission seen in 
all quasars. At the highest inclinations, the 
cooler, lower ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines lower ionization species such as \mg,
\al and Fe~\textsc{ii}, as well as strong Lyman continuum absorption.
The potential links to LoBALQSOs and FeLoBALQSOs are discussed in section 2.4.
% Figure 5 also shows the spectrum into the optical,
% showing that the outflow produces signiciant optical emission lines
% as well as having a noticeable effect on the continuum shape. This
% is particularly true around the Balmer jump at $3646\AA$ and is due to
% a prominent wind-formed recombination continuum.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/alpha_ox_both.png}
\caption
{
X-ray properties of the H13 and clumped model (text filled 
squares), plotted against monochromatic luminosity 
at 2500\AA. Also plotted are the samples considered by
Saez et al. 2012 on a similar plot; The COMBI-7 AGN and
the BQS samples Steffen et al. (2006) and the Saez et al. (2012) sample of BALQSOs.
The dotted lines show the best fit relations for non-BALQSOs from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure*} %fullpage

\begin{figure*}
% \centering
% \includegraphics[width=1.0\textwidth]{figures/sed_all_balqsos.png}
% \caption
% {
% Broadband SEDs compared to IR and X-ray SEDs for selected BALQSOs 
% from Grupe \& Nousek (2015).
% }
% \label{fig:xray}
\end{figure*}

Figure~\ref{fig:xray} shows the emergent
monochromatic luminosity ($L_\nu$) at 2~keV and $\alpha_{OX}$ plotted against 
$L_\nu$ at $2500$\AA\ for a number of different viewing angles in our model.
$\alpha_{OX}$ is a spectral index from near UV to X-rays defined by

\begin{equation}
\alpha_{OX}=0.3838\log\left(\frac{L_{\nu}(2~keV)}{L_{\nu}(2500~\mbox{\scriptsize{\AA}})}\right),
\end{equation}

which effectively represents a near-UV to X-ray flux ratio and a measure of X-ray 
weakness. The properties are calculated from the synthetic spectra and thus include
the effects of wind reprocessing and attenuation. In addition to model outputs,
we also show the BALQSO sample of Saez et al. (2012) and luminous AGN and quasar
samples from Steffen et al. (2006). The best fit relations from Steffen et al. (2006)
are also shown.
For low inclination, `quasar-like' viewing angles,
we now show excellent agreement with AGN samples. The trend with inclination
in our models is caused by a combination of disc foreshortening/limb-darkening 
(resulting in a lower $L_{2500}$ for higher inclinations) and the fact that the disk 
is opaque and thus the X-ray source subtends a smaller solid angle at high inclinations
(resulting in a lower $L_{2keV}$ for higher inclinations).

Although the low inclination, `BALQSO-like' viewing angles show moderate agreement with the data,
it appears that our models tend to over-predict the emergent X-ray luminosity at BAL angles, 
although we are limited by poor statistics. 
% It is possible that this is due to our wind being overly optically thick to 
% electron scattering at angles which look through the wind. 
It could also be that the isotropic X-ray source assumption 
is incorrect and BALQSOs are {\em intrinsically} 
X-ray weak \citep[e.g.][]{morabito2013}. 
an anisotropic X-ray source might result in a lower clumping factor being
required in our model.
Nevertheless, our input X-ray spectrum
now reproduces the X-ray properties of a luminous quasar,
and at least some BAL angles match the .

Typically, BALQSOs show strong X-ray absorption with columns 
of $N_H\sim10^{23}~\rm{cm^{-2}}$ 
\citep{green1996,mathur2000,green2001,grupemathur2003}.
This is often cited as evidence that the BAL outflow is shielded from
the X-ray source, especially as sources with strong X-ray absorption tend
to exhibit deep BAL troughs and high outflow velocities 
\citep{brandt2000,laorbrandt2002,gallagher2006}.
Our results tend to imply that a clumpy BAL outflow
itself can be responsible for the strong X-ray absorption, 
and supports Hamann et al.'s (2013) suggestion that 
this explains the weaker X-ray absorption in mini-BALs 
compared to BALQSOs.

% We can also examine the emergent X-ray spectra from our model. 
% Our code is not yet optimized for producing X-ray spectra, as we do not yet 
% include all lines and levels of e.g. Fe or macro-atom treatments of 
% higher order ions (cf. Sim et al. 2008, 2010). 
% However, the bound-free opacity sources are complete, and comparisons of the shape
% of the heavily absorbed spectra can be useful.
% Figure 6 shows the broadband SED from the model, compared to a number of BALQSO spectra
% from \cite{grupenousek2015} - note that lower mass and super-Eddington sources
% are excluded.

@@

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/c4_angles.png}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for wind angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure*} %fullpage

At certain sightlines, our model now produces blue-shifted BALs in \al\ and \mg --
the absorption lines seen in LoBALQSOs, and we even see absorption in Fe~\textsc{ii}
at the highest inclinations. Line profiles in velocity space 
for \civ, \al\ and \mg, are shown in figure~\ref{fig:lobal} for a range
of BALQSO viewing angles. We confirm the behaviour expected from 
a unification model such as Elvis (2000), in which ionization stratification
explains the incidence of LoBALQSOs. We find, that the lower
ionization material is intercepted by a smaller family of sightlines as shown by 
figures~\ref{fig:wind} and \ref{fig:lobal}.
There is also a correlation between the strength of LoBAL features
and the amount of continuum attenuation at that sightline. 
This offers a potential explanation for the rarity of LoBAL and
FeLoBAL quasars (but see also section~\ref{sec:balfrac}).

% These line profiles also indicate a potential problem with our model. 
% \cite{odowd2015} find that the LoBAL profiles in H1413+117 have a 
% higher velocity onset than the higher ionization BALs, suggesting that the
% wind becomes less ionized as radius increases. 
% We find the opposite trend in velocity onsets. In addition, examination of 
% figure~\ref{fig:wind} shows that ionization parameter decreases with radius, 
% because the decrease in density wins over geometric dilution of the radiation 
% field. While the \cite{odowd2015} hypothesis comes from one object,
% it is clear that {\em at least some} BALQSOs must become less ionized at 
% higher velocities, contrary to what one expects from our modelling.

@@

\section{Discussion: The Disc SED}

When comparing BALQSO and quasar composites, it is apparent
that they possess remarkably similar line strengths and widths 
\citep[e.g.][]{reichard2003}.
This presents a challenge to our model, as well as the geometric 
unification picture in general.
Limb darkening and foreshortening causes the disc continuum emission to be 
strongly anisotropic. Coupled with attenuation from the wind, 
this has the effect of enhancing the 
line-to-continuum ratios at high inclinations, to the extent
that the \civ\ line can have an equivalent widths of ?? at $80^\circ$
compared to ?? at $20^\circ$. To construct a scenario where emission 
line equivalent widths are comparable at all inclinations requires 
significant fine-tuning when considering a classical thin disc. 
An alternative, simpler solution is that the emergent continuum is roughly isotropic.
This could potentially be achieved by wind reprocessing or a more realistic, GR treatment
of the accretion disc. We briefly discuss both these possibilities.

If the majority of continuum radiation was reprocessed by nearby photoionized plasma (such 
as the outflow itself), then this would result in a roughly
isotropic radiation field. There are a number of appealing aspects to this scenario.
First, it may go some way to explaining recent microlensing (REFs) and timing (REFs)
observations which suggest that accretion disc sizes are a factor of $\sim3$ 
larger than one expects from a Shakura \& Sunyaev (1973) thin disc. 
Second, it implies a wide angle, large covering factor wind in order to 
intercept most of the disc continuum. This would make the outflow 
an effective feedback mechanism \citep[e.g.][]{borguet2012}, 
and would imply that BALQSO outflows
have a similar geometry to the ultra-fast outflows seen in nearby AGN 
\citep[e.g.][]{reeves2003,poundsreeves2009,tombesi2010a}.
We note that spectropolarimetry of BALQSOs shows that the light from BAL troughs is $\sim10\%$ 
linearly polarised, but the continuum generally has negligible polarisation \citep{lamy2000}.
This suggests that the continuum of BALQSOs cannot be significantly enhanced by
{\em single} scattering into the line of sight, but does not rule out 
{\em multiple} scatterings leading to a more isotropic continuum.

@@

General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}.
To generate GR disc spectra, we use the code \agn\ \citep{hubeny2000,davishubeny2006,davis2007}. 
The output flux at three different wavelengths
as a function of inclination for an \agn\ model with the same disc and BH parameters
as our clumpy wind model is shown in figure~\ref{fig:f2000}.
The effects of GR on an AGN disc are much less extreme 
in the UV portion of the spectrum than the calculations
by \cite{zhang1997} for the X-rays in X-ray binaries.
As a result, the disc emission is still strongly anisotropic.
GR alone therefore cannot explain the line ratio trends in quasars.

% To examine the effect of GR on our models, we carry out the following procedure.
% First, we fit our output spectrum with a polynomial. This continuum fit can 
% be subtracted from the actual spectrum 
% to produce a model of the line emission only. We then calculate the 
% `wind continuum factor', by comparing the model fit with the emergent 
% disc spectrum without a wind. This factor gives us an approximate way to quantify the 
% effect of the wind on a disc continuum. It will generally be $>1$ at high inclinations,
% and $<1$ at low inclinations. This calculates emergent accretion disc spectra from an AGN disc atmosphere
% including relativistic effects. Finally, produce our final synthetic spectrum 
% by convolving this the line emission spectrum with the \agn\ model for the model parameters.

% Figure 10 shows the resultant spectrum. Including GR clearly helps

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/agnspec.png}
\caption
{
$F_\lambda$ for three different wavelengths
as a function of inclination from \agn\ models, 
compared to a classical AD. 
The \agn\ models are computed for Kerr and Scharzschild
BHs with the same $M_{BH}$ and $\dot{m}$ as our model.
}
\label{fig:f2000}
\end{figure}

% \begin{figure*} %fullpage
% \centering
% \includegraphics[width=1.0\textwidth]{figures/adjusted_withbals.png}
% \caption
% {
% $F_\lambda$ normalised to $F_{2000}$. {\bf Again, this is a placeholder- but I'm thinking some kind of comparison to composite including the adjusted continuum, showing that we can't get it exactly right}.
% }
% \label{fig:finalcomp}
% \end{figure*} %fullpage

\subsection{The BALQSO fraction and wind covering factor}
\label{sec:balfrac}

The disc SED has a profound effect on the intrinsic BAL fraction
inferred from flux-limited samples. This in turn impacts on 
the geometry and covering factor assumed for BALQSO and unified wind models,
and even effects the efficiency of feedback.
\cite{krolik1998} noted that the effect of limb darkening and foreshortening
would result in a signicantly underestimated covering factor for BALQSO winds.
The effect is complicated further by the quasar luminosity function -- if 
BALQSOs have lower observed luminosities for given accretion disc properties 
then they may be drawn from a different bin in the quasar luminosity function 
\citep{goodrich1997}. This will also have a similar impact
on the LoBALQSOs and FeLoBALQSOs sub-populations.
The complex (and as yet unknown) selection effects are understandably 
not taken into account by current estimates of the intrinsic BALQSO fraction 
\citep{weymann1991, reichard2003, knigge2008, turnermiller2009, allen2011}.

The way forward is somewhat unclear. Given the lack of available observational 
constraints, models cannot readily be used to inform the selection criteria when 
calculating the BALQSO fraction. There are two main possibilities:
First, the disc emission is roughly isotropic, in which case the BAL fraction
estimates should be a good estimate of the covering factor of the outflow,
but the question about how this isotropy is achieved remains outstanding.
Or second, the disc emission is strongly anisotropic, which follows more
naturally from a disc geometry but has large implications
for the inferred covering factors, as well as leaving the similar line strengths
in BAL and non-BAL quasar compoosites unexplained (see previous sections).
Understanding the true nature of the disc SED, including the
angular distribution of radiation, is clearly crucial in order
to accurately estimate the true covering factor of BALQSO winds.

@@

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/f2000_m9.png}
% \caption
% {
% $F_\lambda$ at $2000$~\AA\ as a function of inclination from
% \agn\ models. Spectra are computed for BH of $10^9~M_\odot$ with
% a number of different BH spins and Eddington fractions. The black line
% shows a standard multi-temperature blackbody AD model.
% }
% \label{fig:alpha_ox}
% \end{figure}

@@

\section{Summary}

@@

\begin{enumerate}
\item We have introduced a first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We have shown that the degree of ionization stratification
in the model is sufficient that LoBAL line profiles
are seen at a subset of viewing angles, and Fe~\textsc{ii}
absorption is seen at particularly high inclinations.
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models increase with inclination.
BAL and non-BAL quasar composites have comparable EWs, so our model
fails to reproduce this behaviour.
This is due to a fundamental constraint discussed further in section 5. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is simply not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry and size.
\smallskip
\item We have examined the effect of GR on our disc SED, using the disc atmosphere
and GR ray-tracing code \agn. While including GR effects
does cause the disc SED to become slightly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We briefly discuss other solutions.
\end{enumerate}
Our work confirms a number of expected outcomes from a geometric unification 
model, and suggests that a simple biconical geometry such as this can come close to 
explaining much of the  phenomenology of quasars. Nevertheless, our conclusions pose 
a clear challenge to the current disc wind unification picture.
